---
title: "Loch Eriboll seabed map"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

#setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#required packages

 library(sf)
 library(RColorBrewer)
 library(fields)
 library(terra)
 library(gstat)
 library(randomForest)
 library(mgcv)


########## import data ########

### Shape files inData/GIS folder 
#                       - shore <- shoreline.shp




### rasters
# in Data/GIS folder- depth <- bathy.grd
#                   -  slo <- slope.grd
#                   -  asp <- aspect.grd


# Data file in data data folder
#                   - dat<- Loch_Eriboll_substratum_combine_cob_boulder.csv



# Data file in Data/GIS/Substrata_Output_from_script_1 folder

# substrata
# mud.cov.r<- mud_se_pred.grd
# 
# muddy_sand.cov.r<- muddy_s_se_pred.grd
# 
# sand.cov.r<- sand_se_pred.grd
# 
# 
# gravel.cov.r<- gravel_se_pred.grd
# 
# cob_boulder.cov.r<- cob_se_boul_pred.grd
# 
# rock.cov.r<- rock_se_pred.grd
# 
############################################
# Read in rasters of OK and RF for supplemental material
####################################

# Cov for GAM model
# mud.rf<- mud_pred_rf.grd
# 
# muddy_sand.rf<- muddy_s_rf_pred.grd
# 
# sand.rf<- sand_rf_pred.grd
# 
# 
# gravel.rf<- gravel_rf_pred.grd
# 
# cob_boulder.rf<- cob_boul_rf_pred.grd
# 
# rock.rf<- rock_rf_pred.grd


# RF
# mud.rf<- mud_pred_rf.grd
# 
# muddy_sand.rf<- muddy_s_rf_pred.grd
# 
# sand.rf<- sand_rf_pred.grd
# 
# 
# gravel.rf<- gravel_rf_pred.grd
# 
# cob_boulder.rf<- cob_boul_rf_pred.grd
# 
# rock.rf<- rock_rf_pred.grd


# Krig
# mud.k<- mud_pred_k.grd
# 
# muddy_sand.k<- muddy_s_k_pred.grd
# 
# sand.k<- sand_k_pred.grd
# 
# 
# gravel.k<- gravel_k_pred.grd
# 
# cob_boulder.k<- cob_boul_k_pred.grd
# 
# rock.k<- rock_k_pred.grd



## data tidy

sub<- subset(dat, dat$Substratum1!='')
sub<- sub[,-12]
sub<- droplevels(sub)




sub$Substratum1<- as.factor(sub$Substratum1)
sub$Substratum2<- as.factor(sub$Substratum2)
sub$Substratum3<- as.factor(sub$Substratum3)




levels(sub$Substratum1)
sub$Substratum1<- factor(sub$Substratum1, levels = c("mud", "muddy_sand","sand", "gravel", "cob_boulder", "rock"))
levels(sub$Substratum1)


levels(sub$Substratum2)
sub$Substratum2<- factor(sub$Substratum2, levels = c("mud", "muddy_sand","sand", "gravel", "cob_boulder", "rock"))

levels(sub$Substratum3)
sub$Substratum3<- factor(sub$Substratum3, levels = c("sand", "cob_boulder"))



## make an spatial from the data
sub.sf<- st_as_sf(x = sub, 
                        coords = c("x", "y"), remove=FALSE,
                        crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")

sub.sf<- st_transform(sub.sf, crs = 'EPSG:27700')
sub.sf<- cbind(sub.sf, st_coordinates(sub.sf))

#make this back into data frame so we have national grid x and y
sub<- as.data.frame(sub.sf)

#colours for prediction maps
couls<- rev(brewer.pal(11,'Spectral'))
couls<- c( '#FFFFFF', couls)

prob.heat<- colorRampPalette(couls)

depth.heat<- colorRampPalette(rev(brewer.pal(9,'Blues')[2:9]))


rf.cols<- c(brewer.pal(11,'Spectral')[11], brewer.pal(11,'Spectral')[1])


```


#Background map
```{r}


hill.col<- colorRampPalette(c('floralwhite','antiquewhite1', 'antiquewhite2', 'antiquewhite3', 'antiquewhite4'), bias=0.5)


depth.heat<- colorRampPalette(rev(brewer.pal(9,'Blues')[2:9]))

hill<- shade(slo, asp, angle=55, direction=100, normalize =T)

plot(hill, col=hill.col(1000), legend=F)


plot(depth, add=T, legend=F, col=depth.heat(1000))
plot(st_geometry(sub.sf), add=T, pch=19, col='dodgerblue3', cex=0.5)



```



# fig 1 sample location maps
```{r Spatial data, include=FALSE}



par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F, axes=NULL)

plot(depth, add=T, legend=F, col=depth.heat(1000))
plot(shore, add=T, col='black', lwd=0.8)
plot(st_geometry(sub.sf), add=T, pch=19, col='black', cex=0.7)


```


# Substrata vecrors tidy 
```{r}



plot(hill, col=hill.col(1000), legend=F, add=T)
plot(depth, add=T, col=depth.heat(1000), legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)



plot(sub.sf, pch=19, add=T, col=
       brewer.pal(8,"Dark2")[sub.sf$Substratum1], cex=2)

sub.sf2<- sub.sf[,10]

 
```

#Fig 3 Plot
```{r}

# set up some colour options
 test.cols <- c(
     rgb(253,192,134, maxColorValue = 255),
     rgb(127,201,127, maxColorValue = 255),
     rgb(255,255,153, maxColorValue = 255),
     rgb(240,2,127, maxColorValue = 255),
     rgb(190,174,212, maxColorValue = 255),
     rgb(56,108,176, maxColorValue = 255)
     )


test.cols2<- c(
            rgb(255,127,0, maxColorValue = 255),
             rgb(77,175,74, maxColorValue = 255),
            rgb(255,255,51, maxColorValue = 255),
            rgb(228,26,28, maxColorValue = 255),
           
            rgb(152,78,163, maxColorValue = 255),
            rgb(55,126,184, maxColorValue = 255)
            )

par(mar=c(0,0,0,0)+0.1)



plot(hill, col=hill.col(1000), legend=F, axes=NULL)

plot(depth, add=T, legend=F, col=depth.heat(1000))
plot(shore, add=T, col='black', lwd=0.8)

plot(sub.sf, pch=19, add=T, col=
       brewer.pal(8,"Dark2")[c(2:6,8)][sub.sf$Substratum1], cex=1.5)

plot(sub.sf, pch=19, add=T, col=
       brewer.pal(8,"Dark2")[c(2:6,8)][sub.sf$Substratum2], cex=1)

plot(sub.sf, pch=19, add=T, col=
       brewer.pal(8,"Dark2")[c(4,6)][sub.sf$Substratum3], cex=0.5)






## legend

par(mar=c(0,0,0,0)+0.1)
plot(NULL)

leg<- c("mud", "muddy sand", "sand", "gravel", "cobble boulder", "rock" )

legend('center', legend=leg, pch=19, cex=1.3,col= test.cols2, bg='white')



## depth legend
 
plot(NULL)


par(mar=c(0,0,0,0)+0.1)

image.plot((depth*-1), horizontal=F, smallplot= c(0.6, 0.63,0.1,0.9), col = depth.heat(1000), legend.only=TRUE,  axis.args = list(cex.axis = 1,
at = c(0, 20, 40, 60, 80, 100)+15, labels =rev(c("0","20","40","60","80","100")), 
tcl=0.3, lwd=0.5, hadj=0.5), axes=F, zlim=c(0, 115))


```

#Mud

## Mud model predictions 
```{r}



mud.r<- mud.r



par(mar=c(0,0,0,0)+0.1)

plot(sub.sf, col='white')
plot(hill.r, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
plot(shore, col='black', add=T, lwd=1)

plot(mud.r, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000) )

plot(sub.sf, pch=19, add=T, col=
       'black', cex=0.9)

plot(sub.sf[sub.sf$Substratum1 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)



#mud legend


par(mar=c(0,0,0,0)+0.1)
plot(NULL)

legend('center', bg='white',pch=19, cex=1.3, legend=leg, col=c( test.cols2[1], 'black', 'black', 'black', 'black', 'black'))






```

#heat map legend for all substrata
```{r}

par(mar=c(0,0,0,0)+0.1)

image.plot((mud.r), horizontal=F, smallplot= c(0.6, 0.63,0.1,0.9), col = prob.heat(1000), legend.only=TRUE,  axis.args = list(cex.axis = 1,
at = c(0, 22, 50, 75, 100), labels =c("0","0.25","0.5","0.75","1.0"), 
tcl=0.3, lwd=0.5, hadj=0.5), axes=F, zlim=c(0, 100))


```

## Mud SE
```{r}



par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(depth, add=T, col='white', legend=FALSE, axes=FALSE)


plot(mud.cov.r, add=T, legend=FALSE, axes=FALSE, col=rev(terrain.colors(1000)) )
plot(shore, col='black', add=T, lwd=1)
plot(sub.sf[1], pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "mud" ,], pch=19, add=T, col=
       brewer.pal(8,"Dark2")[2], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "mud" ,], pch=19, add=T, col=
       brewer.pal(8,"Dark2")[2], cex=1.1)

```


##SE heat map legend
```{r}

plot(NULL)


par(mar=c(0,0,0,0)+0.1)

image.plot((mud.cov.r), horizontal=F, smallplot= c(0.6, 0.63,0.1,0.9), col = rev(terrain.colors(1000)), legend.only=TRUE,  axis.args = list(cex.axis = 1,
at = c(0, 22, 50, 75, 100), labels =c("0","0.25","0.5","0.75","1.0"), 
tcl=0.3, lwd=0.5, hadj=0.5), axes=F, zlim=c(0, 100))




```



## Mud RF prediction
```{r}



par(mar=c(0,0,0,0)+0.1)


plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(depth, col='white', add=T, legend=F)
plot(mud.rf, add=T, legend=FALSE, axes=FALSE, col=rf.cols)

plot(shore, col='black', add=T, lwd=1)

plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "mud" ,], pch=19, add=T, col=
      test.cols2[1], cex=1.1)



```


## mud krig predictions
```{r}



par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(mud.k, add=T, legend=FALSE, axes=FALSE, col= prob.heat((1000)))
plot(shore, col='black', add=T, lwd=1)

plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)



```



# Muddy_sand

## Muddy_sand model predictions
```{r}


par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)

plot(muddy_s.r, col=prob.heat(1000), add=T, legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)

plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

# legend

 par(mar=c(0,0,0,0)+0.1)
 plot(NULL)
 
 legend('center', bg='white',pch=19, cex=1.3, legend=leg, col=c('1', test.cols2[2], '1', '1', '1','1'))







```


## Muddy sand SE
```{r}


par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(depth, add=T, col='white', legend=FALSE, axes=FALSE)


plot(muddy_sand.cov.r, add=T, legend=FALSE, axes=FALSE, col=rev(terrain.colors(1000)) )
plot(shore, col='black', add=T, lwd=1)
plot(sub.sf[1], pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)



```

## Muddy_sand RF prediction
```{r}



par(mar=c(0,0,0,0)+0.1)


plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(depth, col='white', add=T, legend=F)

plot(muddy_s.rf, add=T, legend=FALSE, axes=FALSE, col=rf.cols)
plot(shore, col='black', add=T, lwd=1)


plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)



```


##muddy_sand krig
```{r}


par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(muddy_s_k, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)


```


# sand

## sand model predictions
```{r}




par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)


plot(sand.r, col = prob.heat(1000), add=T, legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)


plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "sand" ,], pch=19, add=T, col=
       test.cols2[3], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

plot(sub.sf[sub.sf$Substratum3 == "sand" ,], pch=19, add=T, col=
      test.cols2[3], cex=1.1)

# legend

par(mar=c(0,0,0,0)+0.1)
plot(NULL)

legend('center', bg='white', pch=19, cex=1.3, legend=leg, col=c('1', '1',  test.cols2[3], '1', '1', '1'))





```


## Sand SE
```{r}




par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(depth, add=T, col='white', legend=FALSE, axes=FALSE)


plot(sand.cov.r, add=T, legend=FALSE, axes=FALSE, col=rev(terrain.colors(1000)) )
plot(shore, col='black', add=T, lwd=1)
plot(sub.sf[1], pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "sand" ,], pch=19, add=T, col=
       test.cols2[3], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

plot(sub.sf[sub.sf$Substratum3 == "sand" ,], pch=19, add=T, col=
      test.cols2[3], cex=1.1)



```



## sand RF prediction
```{r}




par(mar=c(0,0,0,0)+0.1)


plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(depth, col='white', add=T, legend=F)

plot(sand.rf, add=T, legend=FALSE, axes=FALSE,col=rf.cols)
plot(shore, col='black', add=T, lwd=1)

plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

plot(sub.sf[sub.sf$Substratum3 == "sand" ,], pch=19, add=T, col=
      test.cols2[3], cex=1.1)





```


##sand krig
```{r}


par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(sand_k, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

plot(sub.sf[sub.sf$Substratum3 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)




```



# gravel


## gravel model predictions
```{r}



par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)

plot(gravel.r, col = prob.heat(1000), add=T, legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)


plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

# legend

par(mar=c(0,0,0,0)+0.1)
plot(NULL)


legend('center',bg='white', pch=19, cex=1.3, legend=leg, col=c('1','1','1', test.cols2[4], '1', '1'))



```


## gravel SE
```{r}



par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(depth, add=T, col='white', legend=FALSE, axes=FALSE)


plot(gravel.cov.r, add=T, legend=FALSE, axes=FALSE, col=rev(terrain.colors(1000)) )
plot(shore, col='black', add=T, lwd=1)
plot(sub.sf[1], pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)



```



## gravel RF prediction
```{r}
 


par(mar=c(0,0,0,0)+0.1)


plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(depth, col='white', add=T, legend=F)

plot(gravel.rf, add=T, legend=FALSE, axes=FALSE, col=rf.cols)
plot(shore, col='black', add=T, lwd=1)

plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)



```


##gravel krig prdiction
```{r}



par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(gravel.k, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

#legend

legend('bottomright', legend=leg, fill=c('black', 'black', 'black', test.cols2[4], 'black', 'black'))




```



#Cobble-boulder


## Cob_boulder model predictions
```{r}



par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)

plot(cob_boulder.r, col = prob.heat(1000), add=T, legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)


plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)


plot(sub.sf[sub.sf$Substratum2 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)

plot(sub.sf[sub.sf$Substratum3 == "cob_boulder" ,], pch=19, add=T, col=
     test.cols2[5], cex=1.1)

#legend



par(mar=c(0,0,0,0)+0.1)
plot(NULL)

legend('center',bg='white', pch=19, cex=1.3, legend=leg, col=c('1','1','1','1', test.cols2[5],'1'))





```


## cob_boulder SE
```{r}


par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(depth, add=T, col='white', legend=FALSE, axes=FALSE)


plot(cob_boulder.cov.r,add=T, legend=FALSE, axes=FALSE, col=rev(terrain.colors(1000)) )
plot(shore, col='black', add=T, lwd=1)
plot(sub.sf[1], pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)


plot(sub.sf[sub.sf$Substratum2 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)

plot(sub.sf[sub.sf$Substratum3 == "cob_boulder" ,], pch=19, add=T, col=
     test.cols2[5], cex=1.1)



```








## Cob-boulder RF prediction
```{r}



par(mar=c(0,0,0,0)+0.1)


plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(depth, add=T, col='white', legend=FALSE, axes=FALSE)


plot(cob_boulder.cov.r,add=T, legend=FALSE, axes=FALSE, col=rev(terrain.colors(1000)) )
plot(shore, col='black', add=T, lwd=1)
plot(sub.sf[1], pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)


plot(sub.sf[sub.sf$Substratum2 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)

plot(sub.sf[sub.sf$Substratum3 == "cob_boulder" ,], pch=19, add=T, col=
     test.cols2[5], cex=1.1)m3 == "cob_boulder" ,], pch=19, add=T, col=
     test.cols2[5], cex=1.1)



```


##Cob_boulder krig
```{r}

 
par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(cob_boulder.k, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)

plot(sub.sf[sub.sf$Substratum3 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)



```


# rock


## rock model predictions
```{r}





par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)

plot(rock.r, col = prob.heat(1000), add=T, legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)


plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)



# legend
par(mar=c(0,0,0,0)+0.1)
plot(NULL)


legend('center', bg='white', pch=19, cex=1.3, legend=leg, col= c('1','1','1','1','1', test.cols2[6]))





```


## rock SE
```{r}



par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(depth, add=T, col='white', legend=FALSE, axes=FALSE)


plot(rock.cov.r, add=T, legend=FALSE, axes=FALSE, col=rev(terrain.colors(1000)) )
plot(shore, col='black', add=T, lwd=1)
plot(sub.sf[1], pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)



```





## rock RF prediction
```{r}



par(mar=c(0,0,0,0)+0.1)


plot(hill, col=hill.col(1000), legend=F,  axes=FALSE, box=F)
plot(depth, col='white', add=T, legend=F)

plot(rock.rf, add=T, legend=FALSE, axes=FALSE, col=rf.cols)
plot(shore, col='black', add=T, lwd=1)

plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)



```


##rock krig
```{r}

 
par(mar=c(0,0,0,0)+0.1)

plot(hill, col=hill.col(1000),legend=F,  axes=FALSE, box=F)
plot(rock.k, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.sf, pch=19, add=T, col=
       'black', cex=1)

plot(sub.sf[sub.sf$Substratum1 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)

plot(sub.sf[sub.sf$Substratum2 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)


```



# model plots heatmap legend
```{r}


image.plot((mud.r), horizontal=T, smallplot= c(0.1, 0.9,0.6,0.63), col = rev(terrain.colors(1000)), legend.only=TRUE,  axis.args = list(cex.axis = 1,
at = c(0, 50, 100), labels =c("0","0.5","1.0"), 
tcl=0.3, lwd=0.5, hadj=0.5), axes=F, zlim=c(0, 100))



```


## Area calculations set up
```{r}

# use frequency of probabilities to trim each raster to display reasonable estimates and calculate areas
# vary trim to to change level of probability of presence and display plot as required
density(mud.r)
density(muddy_sand.r)
density(sand.r)
density(gravel.r)
density(cob_boulder.r)
density(rock.r)


mud.r<- mud.r
muddy_sand.r<- muddy_sand.r
sand.r<- sand.r
gravel.r<- gravel.r
cob_boulder.r<- cob_boulder.r
rock.r<- rock.r

# E.G of setting substrata boundaries 
mud.r [mud.r< 0.51] <- NA
muddy_sand.r [muddy_sand.r< 0.51] <- NA
sand.r [sand.r< 0.11] <- NA
gravel.r [gravel.r< 0.05] <- NA
cob_boulder.r [cob_boulder.r< 0.11] <- NA
rock.r [rock.r< 0.51] <- NA



```


## area calculations
```{r}

########  code written before switch to sf and terra - code written for sp and raster
# overall area
plot(depth, colNA="dodgerblue")
cell.No<-cellStats(depth, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(depth)[1]*res(depth)[2])
round(raster_area/(1000*1000), 2)

mud.r<- mud.r
muddy_sand.r<- muddy_sand.r
sand.r<- sand.r
gravel.r<- gravel.r
cob_boulder.r<- cob_boulder.r
rock.r<- rock.r


hist(mud.r)
mud.r [mud.r<= 0.1] <- NA
density(mud.r)


###area calc
plot(mud.r, colNA="dodgerblue")

cell.No<-cellStats(mud.r, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(mud.r)[1]*res(mud.r)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)

#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

density(muddy_sand.r)
muddy_sand.r [muddy_sand.r<= 0.1] <- NA
density(muddy_sand.r)

###area calc
plot(muddy_sand.r, colNA="dodgerblue")

cell.No<-cellStats(muddy_sand.r, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(muddy_sand.r)[1]*res(muddy_sand.r)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)

#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

density(sand.r)
sand.r [sand.r<= 0.1] <- NA
density(sand.r)

###area calc
plot(sand.r, colNA="dodgerblue")

cell.No<-cellStats(sand.r, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(sand.r)[1]*res(sand.r)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)
#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)



density (gravel.r)
gravel.r [gravel.r <= 0.1] <- NA
density(gravel.r)

###area calc
plot(gravel.r, colNA="dodgerblue")

cell.No<-cellStats(gravel.r, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(gravel.r)[1]*res(gravel.r)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)
#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

density(cob_boulder.r)
cob_boulder.r [cob_boulder.r<= 0.1] <- NA
density(cob_boulder.r)

###area calc
plot(cob_boulder.r, colNA="dodgerblue")

cell.No<-cellStats(cob_boulder.r, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(cob_boulder.r)[1]*res(cob_boulder.r)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)
#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

density(rock.r)
rock.r [rock.r<= 0.1] <- NA
density(rock.r)

###area calc
plot(rock.r, colNA="dodgerblue")

cell.No<-cellStats(rock.r, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(rock.r)[1]*res(rock.r)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)
#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

```
