---
title: "Loch Eriboll seabed map"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

#setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#required packages

 library(sp)
 library(RColorBrewer)
 library(fields)
 library(maptools)
 library(raster)
 library(rgdal)
 library(gstat)
 library(flux)
 library(randomForest)
 library(mgcv)


########## import data ########

### Shape files in GIS data folder 
#                       - shore <- shoreline.shp


# map extent
new.ext<- extent(c(xmin=238800, xmax=252000, ymin=953700, ymax=969700))
new.ext

shore<- crop(shore, new.ext)

### rasters
# in GIS data folder- depth <- bathy.grd
#                   -  slo <- slope.grd
#                   -  asp <- aspect.grd


# Data file in GIS data folder
#                   - dat<- Loch_Eriboll_substratum_combine_cob_boulder.csv



# Data file in GIS/Substrata_Output_from_script_1 folder

# substrata
# mud.r<- mud_pred.grd
# 
# muddy_sand.r<- muddy_s_pred.grd
# 
# sand.r<- sand_pred.grd
# 
# 
# gravel.r<- gravel_pred.grd
# 
# cob_boulder.r<- cob_boul_pred.grd
# 
# rock.r<- rock_pred.grd
# 

#########################
# Read in rasters of OK and RF for supplemental material
####################################

# RF
# mud.rf<- mud_pred_rf.grd
# 
# muddy_sand.rf<- muddy_s_rf_pred.grd
# 
# sand.rf<- sand_rf_pred.grd
# 
# 
# gravel.rf<- gravel_rf_pred.grd
# 
# cob_boulder.rf<- cob_boul_rf_pred.grd
# 
# rock.rf<- rock_rf_pred.grd


# Krig
# mud.k<- mud_pred_k.grd
# 
# muddy_sand.k<- muddy_s_k_pred.grd
# 
# sand.k<- sand_k_pred.grd
# 
# 
# gravel.k<- gravel_k_pred.grd
# 
# cob_boulder.k<- cob_boul_k_pred.grd
# 
# rock.k<- rock_k_pred.grd



## data tidy

sub<- subset(dat, dat$Substratum1!='')
sub<- sub[,-12]
sub<- droplevels(sub)




sub$Substratum1<- as.factor(sub$Substratum1)
sub$Substratum2<- as.factor(sub$Substratum2)
sub$Substratum3<- as.factor(sub$Substratum3)




levels(sub$Substratum1)
sub$Substratum1<- factor(sub$Substratum1, levels = c("mud", "muddy_sand","sand", "gravel", "cob_boulder", "rock"))
levels(sub$Substratum1)


levels(sub$Substratum2)
sub$Substratum2<- factor(sub$Substratum2, levels = c("mud", "muddy_sand","sand", "gravel", "cob_boulder", "rock"))

levels(sub$Substratum3)
sub$Substratum3<- factor(sub$Substratum3, levels = c("sand", "cob_boulder"))



table(sub$Substratum1)
table(sub$Substratum2)
table(sub$Substratum3)





################################## coord system to assign 
###

#'#set CRS
CRS1.code <- CRS(SRS_string='EPSG:4326')

#Display the stored CRS using comment()
cat(comment(CRS1.code), "\n")
# Store the wkt in a variable
wkt1 <- comment(CRS1.code)

# Use this to assign the CRS of another sp-object
CRS.wgs84 <- CRS(SRS_string = wkt1)

### projections ####


CRS2.code <- CRS(SRS_string='EPSG:27700')
#Display the stored CRS using comment()
cat(comment(CRS2.code), "\n")
# Store the wkt in a variable
wkt2 <- comment(CRS2.code)

# Use this to assign the CRS of another sp-object
CRS.NG <- CRS(SRS_string = wkt2)


## make an spdf from the data
coord.tmp<- cbind(sub$x, sub$y)
sub.spdf<- SpatialPointsDataFrame(coord.tmp, data=data.frame(sub))
proj4string(sub.spdf) <- CRS.wgs84
sub.spdfT<- spTransform(sub.spdf,  CRS.NG)



```


#Background map
```{r}


hill<- hillShade(slo, asp, angle=55, direction=100, normalize =T)
hill.trim<- hillShade(slo.trim, asp.trim, angle=55, direction=100, normalize =T)


hill.col<- colorRampPalette(c('floralwhite','antiquewhite1', 'antiquewhite2', 'antiquewhite3', 'antiquewhite4'), bias=0.5)


depth.heat<- colorRampPalette(rev(brewer.pal(9,'Blues')[2:9]))



```



# fig 1 sample location maps
```{r Spatial data, include=FALSE}



par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col=NA)

plot(hill.trim, col=hill.col(1000), legend=F, add=T, border=F)
plot(depth, add=T, legend=F, col=depth.heat(1000))
plot(shore, add=T, col='black', lwd=0.8)

plot(sub.spdfT, pch=19, add=T, col='black', cex=0.7)


```


# Substrata vecrors tidy 
```{r}

head(sub.spdfT)
levels(sub.spdfT@data$Substratum1)
levels(sub.spdfT@data$Substratum2)
levels(sub.spdfT@data$Substratum3)


plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(depth, add=T, col=depth.heat(1000), legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)


sub.spdfT@data$Substratum2<- as.factor(sub.spdfT@data$Substratum2)
sub.spdfT@data$Substratum3<- as.factor(sub.spdfT@data$Substratum3)



plot(sub.spdfT, pch=19, add=T, col=
       brewer.pal(8,"Dark2")[sub.spdfT@data$Substratum1], cex=2)

sub.spdfT2<- sub.spdfT[,10]
sub.spdfT3<- sub.spdfT[,11]


sp.na.omit <- function(x, margin=1) {
  if (!inherits(x, "SpatialPointsDataFrame") & !inherits(x, "SpatialPolygonsDataFrame")) 
    stop("MUST BE sp SpatialPointsDataFrame OR SpatialPolygonsDataFrame CLASS OBJECT") 
  na.index <- unique(as.data.frame(which(is.na(x@data),arr.ind=TRUE))[,margin])
    if(margin == 1) {  
      cat("DELETING ROWS: ", na.index, "\n") 
        return( x[-na.index,]  ) 
    }
    if(margin == 2) {  
      cat("DELETING COLUMNS: ", na.index, "\n") 
        return( x[,-na.index]  ) 
    }
 }

# DELETE NA's IN meuse AND SHOW CHANGE IN dim
sub.spdfT2.trim <- sp.na.omit(sub.spdfT2)     
  dim(sub.spdfT2)
    dim(sub.spdfT2.trim)
    
sub.spdfT2.trim@data$Substratum2<- factor(sub.spdfT2.trim@data$Substratum2, levels = c("mud", "muddy_sand","sand", "gravel", "cob_boulder", "rock"))

  
sub.spdfT3.trim <- sp.na.omit(sub.spdfT3)     
  dim(sub.spdfT3)
    dim(sub.spdfT3.trim) 
    
sub.spdfT3.trim@data$Substratum3<- factor(sub.spdfT3.trim@data$Substratum3, levels = c("sand", "cob_boulder"))

  
  
 
```

#Fig 3 Plot
```{r}

# set up some colour options
 test.cols <- c(
     rgb(253,192,134, maxColorValue = 255),
     rgb(127,201,127, maxColorValue = 255),
     rgb(255,255,153, maxColorValue = 255),
     rgb(240,2,127, maxColorValue = 255),
     rgb(190,174,212, maxColorValue = 255),
     rgb(56,108,176, maxColorValue = 255)
     )


test.cols2<- c(
            rgb(255,127,0, maxColorValue = 255),
             rgb(77,175,74, maxColorValue = 255),
            rgb(255,255,51, maxColorValue = 255),
            rgb(228,26,28, maxColorValue = 255),
           
            rgb(152,78,163, maxColorValue = 255),
            rgb(55,126,184, maxColorValue = 255)
            )

par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
plot(depth, add=T, col=depth.heat(1000), legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       test.cols2[sub.spdfT@data$Substratum1], cex=1.2)

plot(sub.spdfT2.trim, pch=19, add=T, col=
       test.cols2[sub.spdfT2.trim@data$Substratum2], cex=0.8)

plot(sub.spdfT3.trim, pch=19, add=T, col=
       test.cols2[c(3,5)][sub.spdfT3.trim@data$Substratum3], cex=0.5)





## legend

par(mar=c(0,0,0,0)+0.1)
plot(sub.spdfT, col=NA)

leg<- c("mud", "muddy sand", "sand", "gravel", "cobble boulder", "rock" )

legend('center', legend=leg, pch=19, cex=1.3,col= test.cols2, bg='white')



## depth legend
 
plot(sub.spdfT, col=NA)


par(mar=c(0,0,0,0)+0.1)

image.plot((depth*-1), horizontal=F, smallplot= c(0.6, 0.63,0.1,0.9), col = depth.heat(1000), legend.only=TRUE,  axis.args = list(cex.axis = 1,
at = c(0, 20, 40, 60, 80, 100)+15, labels =rev(c("0","20","40","60","80","100")), 
tcl=0.3, lwd=0.5, hadj=0.5), axes=F, zlim=c(0, 115))


```

#Mud

## Mud model predictions 
```{r}



mud.trim<- mud.r
mud.trim [mud.trim< 0.1] <- NA
prob.heat<- colorRampPalette(rev(brewer.pal(11,'Spectral')))



par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
plot(shore, col='black', add=T, lwd=1)

plot(mud.trim, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000) )

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=0.9)

plot(sub.spdfT[sub.spdfT$Substratum1 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)



#mud legend


par(mar=c(0,0,0,0)+0.1)
plot(sub.spdfT, col=NA)

legend('center', bg='white',pch=19, cex=1.3, legend=leg, col=c( test.cols2[1], 'black', 'black', 'black', 'black', 'black'))






```

#heat map legend for all substrata
```{r}

par(mar=c(0,0,0,0)+0.1)

image.plot((mud.r), horizontal=F, smallplot= c(0.6, 0.63,0.1,0.9), col = prob.heat(1000), legend.only=TRUE,  axis.args = list(cex.axis = 1,
at = c(0, 22, 50, 75, 100), labels =c("0","0.25","0.5","0.75","1.0"), 
tcl=0.3, lwd=0.5, hadj=0.5), axes=F, zlim=c(0, 100))


```



## Mud RF prediction
```{r}


rf.cols<- c(brewer.pal(11,'Spectral')[11], brewer.pal(11,'Spectral')[1])

par(mar=c(0,0,0,0)+0.1)

mud.rf.trim<- mud.rf
mud.rf.trim [mud.rf.trim<2 ] <- NA

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(mud.rf.trim, add=T, legend=FALSE, axes=FALSE, col=rf.cols)
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "mud" ,], pch=19, add=T, col=
      test.cols2[1], cex=1.1)



```

## mud krig predictions
```{r}


mud.k.trim<- mud.k
mud.k.trim [mud.k.trim< 0.1] <- NA


par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(mud.k.trim, add=T, legend=FALSE, axes=FALSE, col= prob.heat((1000)))
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)



```



# Muddy_sand

## Muddy_sand model predictions
```{r}

#
muddy_s.trim<- muddy_sand.r
muddy_s.trim [muddy_s.trim< 0.1] <- NA

par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
plot(shore, col='black', add=T, lwd=1)

plot(muddy_s.trim, col=prob.heat(1000), add=T, legend=FALSE, axes=FALSE)


plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

# legend

 par(mar=c(0,0,0,0)+0.1)
 plot(sub.spdfT, col=NA)
 
 legend('center', bg='white',pch=19, cex=1.3, legend=leg, col=c('1', test.cols2[2], '1', '1', '1','1'))







```

## Muddy_sand RF prediction
```{r}



muddy_s.rf.trim<- muddy_sand.rf
muddy_s.rf.trim [muddy_s.rf.trim<2 ] <- NA


par(mar=c(0,0,0,0)+0.1)



plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(muddy_s.rf.trim, add=T, legend=FALSE, axes=FALSE, col=rf.cols)
plot(shore, col='black', add=T, lwd=1)


plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)



```


##muddy_sand krig
```{r}



muddy_s_k.trim<- muddy_sand.k
muddy_s_k.trim [muddy_s_k.trim< 0.1] <- NA


par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(muddy_s_k.trim, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)


```


# sand

## sand model predictions
```{r}



sand.trim<- sand.r
sand.trim [sand.trim< 0.1] <- NA



par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
plot(shore, col='black', add=T, lwd=1)


plot(sand.trim, col = prob.heat(1000), add=T, legend=FALSE, axes=FALSE)


plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "sand" ,], pch=19, add=T, col=
       test.cols2[3], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

plot(sub.spdfT3.trim[sub.spdfT3.trim$Substratum3 == "sand" ,], pch=19, add=T, col=
      test.cols2[3], cex=1.1)

# legend

par(mar=c(0,0,0,0)+0.1)
plot(sub.spdfT, col=NA)

legend('center', bg='white', pch=19, cex=1.3, legend=leg, col=c('1', '1',  test.cols2[3], '1', '1', '1'))





```

## sand RF prediction
```{r}



sand.rf.trim<- sand.rf
sand.rf.trim [sand.rf.trim<2 ] <- NA


par(mar=c(0,0,0,0)+0.1)


plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(sand.rf.trim, add=T, legend=FALSE, axes=FALSE,col=rf.cols)
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

plot(sub.spdfT3.trim[sub.spdfT3.trim$Substratum3 == "sand" ,], pch=19, add=T, col=
      test.cols2[3], cex=1.1)




```


##sand krig
```{r}



sand_k.trim<- sand.k
sand_k.trim [sand_k.trim< 0.1] <- NA


par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(sand_k.trim, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)




```



# gravel


## gravel model predictions
```{r}




gravel.trim<- gravel.r
gravel.trim [gravel.trim< 0.1] <- NA


par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
plot(shore, col='black', add=T, lwd=1)


plot(gravel.trim, col = prob.heat(1000), add=T, legend=FALSE, axes=FALSE)


plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

# legend

par(mar=c(0,0,0,0)+0.1)
plot(sub.spdfT, col=NA)


legend('center',bg='white', pch=19, cex=1.3, legend=leg, col=c('1','1','1', test.cols2[4], '1', '1'))



```

## gravel RF prediction
```{r}
 

gravel.rf.trim<- gravel.rf
gravel.rf.trim [gravel.rf.trim<2 ] <- NA

par(mar=c(0,0,0,0)+0.1)


plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(gravel.rf.trim, add=T, legend=FALSE, axes=FALSE, col=rf.cols)
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)



```


##gravel krig prdiction
```{r}


gravel.k.trim<- gravel.k
gravel.k.trim [gravel.k.trim< 0.1] <- NA

par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(gravel.k.trim, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

#legend

legend('bottomright', legend=leg, fill=c('black', 'black', 'black', test.cols2[4], 'black', 'black'))




```



#Cobble-boulder


## Cob_boulder model predictions
```{r}


cob_boulder.trim<- cob_boulder.r
cob_boulder.trim [cob_boulder.trim< 0.1] <- NA



par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
plot(shore, col='black', add=T, lwd=1)



plot(cob_boulder.trim, col = prob.heat(1000), add=T, legend=FALSE, axes=FALSE)


plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)


plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)

plot(sub.spdfT3.trim[sub.spdfT3.trim$Substratum3 == "cob_boulder" ,], pch=19, add=T, col=
     test.cols2[5], cex=1.1)

#legend



par(mar=c(0,0,0,0)+0.1)
plot(sub.spdfT, col=NA)

legend('center',bg='white', pch=19, cex=1.3, legend=leg, col=c('1','1','1','1', test.cols2[5],'1'))





```

## Cob-boulder RF prediction
```{r}



cob_boulder.rf.trim<- cob_boulder.rf
cob_boulder.rf.trim [cob_boulder.rf.trim<2 ] <- NA


par(mar=c(0,0,0,0)+0.1)


plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(cob_boulder.rf.trim, add=T, legend=FALSE, axes=FALSE, col=rf.cols)
plot(shore, col='black', add=T, lwd=1)


plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)

plot(sub.spdfT3.trim[sub.spdfT3.trim$Substratum3 == "cob_boulder" ,], pch=19, add=T, col=
     test.cols2[5], cex=1.1)



```


##Cob_boulder krig
```{r}

 

cob_boulder.k.trim<- cob_boulder.k
cob_boulder.k.trim [cob_boulder.k.trim< 0.1] <- NA

par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(cob_boulder.k.trim, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)



```


# rock


## rock model predictions
```{r}



rock.trim<- rock.r
rock.trim [rock.trim< 0.1] <- NA



par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
plot(shore, col='black', add=T, lwd=1)


plot(rock.trim, col = prob.heat(1000), add=T, legend=FALSE, axes=FALSE)


plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)



# legend
par(mar=c(0,0,0,0)+0.1)
plot(sub.spdfT, col=NA)


legend('center', bg='white', pch=19, cex=1.3, legend=leg, col= c('1','1','1','1','1', test.cols2[6]))





```

## rock RF prediction
```{r}

rock.rf.trim<- rock.rf
rock.rf.trim [rock.rf.trim<2 ] <- NA


par(mar=c(0,0,0,0)+0.1)


plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(rock.rf.trim, add=T, legend=FALSE, axes=FALSE, col=rf.cols)
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)



```


##rock krig
```{r}

 
rock.k.trim<- rock.k
rock.k.trim [rock.k.trim< 0.1] <- NA

par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(rock.k.trim, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)


```



# model plots heatmap legend
```{r}


image.plot((mud.r), horizontal=T, smallplot= c(0.1, 0.9,0.6,0.63), col = rev(terrain.colors(1000)), legend.only=TRUE,  axis.args = list(cex.axis = 1,
at = c(0, 50, 100), labels =c("0","0.5","1.0"), 
tcl=0.3, lwd=0.5, hadj=0.5), axes=F, zlim=c(0, 100))



```


## Area calculations set up
```{r}

# use frequency of probabilities to trim each raster to display reasonable estimates and calculate areas
# vary trim to to change level of probability of presence and display plot as required
density(mud.r)
density(muddy_sand.r)
density(sand.r)
density(gravel.r)
density(cob_boulder.r)
density(rock.r)


mud.trim<- mud.r
muddy_sand.trim<- muddy_sand.r
sand.trim<- sand.r
gravel.trim<- gravel.r
cob_boulder.trim<- cob_boulder.r
rock.trim<- rock.r


mud.trim [mud.trim< 0.51] <- NA
muddy_sand.trim [muddy_sand.trim< 0.51] <- NA
sand.trim [sand.trim< 0.11] <- NA
gravel.trim [gravel.trim< 0.05] <- NA
cob_boulder.trim [cob_boulder.trim< 0.11] <- NA
rock.trim [rock.trim< 0.51] <- NA



```


## area calculations
```{r}


# overall area
plot(depth, colNA="dodgerblue")
cell.No<-cellStats(depth, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(depth)[1]*res(depth)[2])
round(raster_area/(1000*1000), 2)

mud.trim<- mud.r
muddy_sand.trim<- muddy_sand.r
sand.trim<- sand.r
gravel.trim<- gravel.r
cob_boulder.trim<- cob_boulder.r
rock.trim<- rock.r


hist(mud.r)
mud.trim [mud.trim<= 0.1] <- NA
density(mud.trim)


###area calc
plot(mud.trim, colNA="dodgerblue")

cell.No<-cellStats(mud.trim, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(mud.trim)[1]*res(mud.trim)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)

#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

density(muddy_sand.r)
muddy_sand.trim [muddy_sand.trim<= 0.1] <- NA
density(muddy_sand.trim)

###area calc
plot(muddy_sand.trim, colNA="dodgerblue")

cell.No<-cellStats(muddy_sand.trim, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(muddy_sand.trim)[1]*res(muddy_sand.trim)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)

#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

density(sand.r)
sand.trim [sand.trim<= 0.1] <- NA
density(sand.trim)

###area calc
plot(sand.trim, colNA="dodgerblue")

cell.No<-cellStats(sand.trim, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(sand.trim)[1]*res(sand.trim)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)
#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)



density (gravel.r)
gravel.trim [gravel.trim <= 0.1] <- NA
density(gravel.trim)

###area calc
plot(gravel.trim, colNA="dodgerblue")

cell.No<-cellStats(gravel.trim, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(gravel.trim)[1]*res(gravel.trim)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)
#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

density(cob_boulder.r)
cob_boulder.trim [cob_boulder.trim<= 0.1] <- NA
density(cob_boulder.trim)

###area calc
plot(cob_boulder.trim, colNA="dodgerblue")

cell.No<-cellStats(cob_boulder.trim, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(cob_boulder.trim)[1]*res(cob_boulder.trim)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)
#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

density(rock.r)
rock.trim [rock.trim<= 0.1] <- NA
density(rock.trim)

###area calc
plot(rock.trim, colNA="dodgerblue")

cell.No<-cellStats(rock.trim, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(rock.trim)[1]*res(rock.trim)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)
#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

```
