---
title: "Loch Eriboll seabed map"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

#To Do


3)  remove CV and modelling code from this script





#setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#setup
#required packages
 library(sp)
 library(RColorBrewer)
 library(fields)
 library(raster)
 library(rgdal)
 library(gstat)
 library(randomForest)
 library(mgcv)


########## import data ########

### Shape files in GIS data folder 
#                       - shore <- shoreline.shp


# map extent
new.ext<- extent(c(xmin=238800, xmax=252000, ymin=953700, ymax=969700))
new.ext

shore<- crop(shore, new.ext)

### rasters
# in GIS data folder- depth <- bathy.grd
#                   -  slo <- slope.grd
#                   -  asp <- aspect.grd


#                   - mud.r<- mud_pred.grd

#                   - muddy_sand.r<- muddy_s_pred.grd
# 
#                   - sand.r<- sand_pred.grd
# 
# 
#                   - gravel.r<- gravel_pred.grd
# 
#                    - cob_boulder.r<- cob_boul_pred.grd
# 
#                   - rock.r<- rock_pred.grd


# start back here continue with same system below as above.##########

#########################
# Read in rasters of OK and RF for supplemental material
####################################

# RF
mud.rf<- raster('F:/SIFT2021/May on/GIS_data/analysis_files/mud_pred_rf.grd')

muddy_sand.rf<- raster('F:/SIFT2021/May on/GIS_data/analysis_files/muddy_s_rf_pred.grd')

sand.rf<- raster('F:/SIFT2021/May on/GIS_data/analysis_files/sand_rf_pred.grd')


gravel.rf<- raster('F:/SIFT2021/May on/GIS_data/analysis_files/gravel_rf_pred.grd')

cob_boulder.rf<- raster('F:/SIFT2021/May on/GIS_data/analysis_files/cob_boul_rf_pred.grd')

rock.rf<- raster('F:/SIFT2021/May on/GIS_data/analysis_files/rock_rf_pred.grd')


# Krig
mud.k<- raster('F:/SIFT2021/May on/GIS_data/analysis_files/mud_pred_k.grd')

muddy_sand.k<- raster('F:/SIFT2021/May on/GIS_data/analysis_files/muddy_s_k_pred.grd')

sand.k<- raster('F:/SIFT2021/May on/GIS_data/analysis_files/sand_k_pred.grd')


gravel.k<- raster('F:/SIFT2021/May on/GIS_data/analysis_files/gravel_k_pred.grd')

cob_boulder.k<- raster('F:/SIFT2021/May on/GIS_data/analysis_files/cob_boul_k_pred.grd')

rock.k<- raster('F:/SIFT2021/May on/GIS_data/analysis_files/rock_k_pred.grd')



### Data

dat<-read.csv("F:/SIFT2021/May on/Loch_Eriboll_substratum_combine_cob_boulder.csv")
head(dat)
str(dat)
names(dat)

sub<- subset(dat, dat$Substratum1!='')
sub<- sub[,-17]
sub<- droplevels(sub)


str(sub)
summary(sub)


sub$Substratum1<- as.factor(sub$Substratum1)
sub$Substratum2<- as.factor(sub$Substratum2)
sub$Substratum3<- as.factor(sub$Substratum3)

levels(sub$Substratum1)
sub$Substratum1<- factor(sub$Substratum1, levels = c("mud", "muddy_sand","sand", "gravel", "cob_boulder", "rock"))
levels(sub$Substratum1)

levels(sub$Substratum2)
sub$Substratum2<- factor(sub$Substratum2, levels = c("mud", "muddy_sand","sand", "gravel", "cob_boulder", "rock"))

levels(sub$Substratum3)
sub$Substratum3<- factor(sub$Substratum3, levels = c("sand", "cob_boulder"))

table(sub$Substratum1)
table(sub$Substratum2)
table(sub$Substratum3)





################################## coord system to assign if needed
###

#'#set CRS
CRS1.code <- CRS(SRS_string='EPSG:4326')

#Display the stored CRS using comment()
cat(comment(CRS1.code), "\n")
# Store the wkt in a variable
wkt1 <- comment(CRS1.code)

# Use this to assign the CRS of another sp-object
CRS.wgs84 <- CRS(SRS_string = wkt1)

### projections ####


CRS2.code <- CRS(SRS_string='EPSG:27700')
#Display the stored CRS using comment()
cat(comment(CRS2.code), "\n")
# Store the wkt in a variable
wkt2 <- comment(CRS2.code)

# Use this to assign the CRS of another sp-object
CRS.NG <- CRS(SRS_string = wkt2)


## make an spdf from the data
coord.tmp<- cbind(sub$x, sub$y)
sub.spdf<- SpatialPointsDataFrame(coord.tmp, data=data.frame(sub))
proj4string(sub.spdf) <- CRS.wgs84
sub.spdfT<- spTransform(sub.spdf,  CRS.NG)


```


#Background map
```{r}


hill<- hillShade(slo, asp, angle=55, direction=100, normalize =T)
hill.trim<- hillShade(slo.trim, asp.trim, angle=55, direction=100, normalize =T)


hill.col<- colorRampPalette(c('floralwhite','antiquewhite1', 'antiquewhite2', 'antiquewhite3', 'antiquewhite4'), bias=0.5)


depth.heat<- colorRampPalette(rev(brewer.pal(9,'Blues')[2:9]))



```



# fig 1 sample location maps
```{r Spatial data, include=FALSE}


#pdf(file="F:/Seabed_map_paper/figures/1_sample_locations.pdf", width=5, height=5)

par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')

plot(hill.trim, col=hill.col(1000), legend=F, add=T, border=F)
plot(depth, add=T, legend=F, col=depth.heat(1000))
plot(shore, add=T, col='black', lwd=0.8)

plot(sub.spdfT, pch=19, add=T, col='black', cex=0.7)

#dev.off()


# plot(sub.spdfT, col='white')
# plot(hill.trim, col=hill.col(1000), legend=F)
# plot(depth, add=T, col=depth.heat(1000), legend=FALSE, axes=FALSE)
# plot(shore, col='black', add=T, lwd=1)
# plot(sub.spdfT, pch=19, add=T, col='dodgerblue3', cex=1)
# 
# plot(mud.r, legend=F)
# plot(shore, col='black', add=T, lwd=1)
# 
# plot(muddy_sand.r, legend=F)
# plot(shore, col='black', add=T, lwd=1)
# 
# 
# plot(sand.r, legend=F)
# plot(shore, col='black', add=T, lwd=1)
# 
# plot(gravel.r, legend=F)
# plot(shore, col='black', add=T, lwd=1)
# 
# 
# plot(cob_boulder.r, legend=F)
# plot(shore, col='black', add=T, lwd=1)
# 
# plot(rock.r, legend=F)
# plot(shore, col='black', add=T, lwd=1)



```


# plot substrata data
 
```{r}

head(sub.spdfT)
levels(sub.spdfT@data$Substratum1)
levels(sub.spdfT@data$Substratum2)
levels(sub.spdfT@data$Substratum3)

#display.brewer.all()

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(depth, add=T, col=depth.heat(1000), legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)


###area calc
#plot(depth, colNA="dodgerblue")

cell.No<-cellStats(depth, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(depth)[1]*res(depth)[2])


total.area <- round(raster_area/(1000*1000), 2)




sub.spdfT@data$Substratum2<- as.factor(sub.spdfT@data$Substratum2)
sub.spdfT@data$Substratum3<- as.factor(sub.spdfT@data$Substratum3)



plot(sub.spdfT, pch=19, add=T, col=
       brewer.pal(8,"Dark2")[sub.spdfT@data$Substratum1], cex=2)

sub.spdfT2<- sub.spdfT[,15]
sub.spdfT3<- sub.spdfT[,16]


### write function to remove '' from spatial data
sp.na.omit <- function(x, margin=1) {
  if (!inherits(x, "SpatialPointsDataFrame") & !inherits(x, "SpatialPolygonsDataFrame")) 
    stop("MUST BE sp SpatialPointsDataFrame OR SpatialPolygonsDataFrame CLASS OBJECT") 
  na.index <- unique(as.data.frame(which(is.na(x@data),arr.ind=TRUE))[,margin])
    if(margin == 1) {  
      cat("DELETING ROWS: ", na.index, "\n") 
        return( x[-na.index,]  ) 
    }
    if(margin == 2) {  
      cat("DELETING COLUMNS: ", na.index, "\n") 
        return( x[,-na.index]  ) 
    }
 }

# DELETE NA's IN meuse AND SHOW CHANGE IN dim
sub.spdfT2.trim <- sp.na.omit(sub.spdfT2)     
  dim(sub.spdfT2)
    dim(sub.spdfT2.trim)
    
sub.spdfT2.trim@data$Substratum2<- factor(sub.spdfT2.trim@data$Substratum2, levels = c("mud", "muddy_sand","sand", "gravel", "cob_boulder", "rock"))


plot(sub.spdfT2, col="red", pch=20)
  plot(sub.spdfT2.trim, col="black", pch=20, add=TRUE)
  
sub.spdfT3.trim <- sp.na.omit(sub.spdfT3)     
  dim(sub.spdfT3)
    dim(sub.spdfT3.trim) 
    
sub.spdfT3.trim@data$Substratum3<- factor(sub.spdfT3.trim@data$Substratum3, levels = c("sand", "cob_boulder"))
    
plot(sub.spdfT3, col="red", pch=20)
  plot(sub.spdfT3.trim, col="black", pch=20, add=TRUE)
  
  
 
```

#Fig 3 Plot
```{r}

 test.cols <- c(
     rgb(253,192,134, maxColorValue = 255),
     rgb(127,201,127, maxColorValue = 255),
     rgb(255,255,153, maxColorValue = 255),
     rgb(240,2,127, maxColorValue = 255),
     rgb(190,174,212, maxColorValue = 255),
     rgb(56,108,176, maxColorValue = 255)
     )


test.cols2<- c(
            rgb(255,127,0, maxColorValue = 255),
             rgb(77,175,74, maxColorValue = 255),
            rgb(255,255,51, maxColorValue = 255),
            rgb(228,26,28, maxColorValue = 255),
           
            rgb(152,78,163, maxColorValue = 255),
            rgb(55,126,184, maxColorValue = 255)
            )

#pdf(file="F:/Seabed_map_paper/figures/3sample_result.pdf", width=5, height=5)
par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
plot(depth, add=T, col=depth.heat(1000), legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       test.cols2[sub.spdfT@data$Substratum1], cex=1.2)

plot(sub.spdfT2.trim, pch=19, add=T, col=
       test.cols2[sub.spdfT2.trim@data$Substratum2], cex=0.8)

plot(sub.spdfT3.trim, pch=19, add=T, col=
       test.cols2[c(3,5)][sub.spdfT3.trim@data$Substratum3], cex=0.5)

#dev.off()




#pdf(file="F:/Seabed_map_paper/Figures/3_sample_legend.pdf",width=4, height=6)


par(mar=c(0,0,0,0)+0.1)
plot(sub.spdfT, col='white')

leg<- c("mud", "muddy sand", "sand", "gravel", "cobble boulder", "rock" )

legend('center', legend=leg, fill= test.cols2, bg='white')



 #dev.off()
 
 
plot(sub.spdfT, col='white')


#pdf(file="F:/Seabed_map_paper/Figures/depth_leg_vert.pdf", width=4.5, height=4.5) 

par(mar=c(0,0,0,0)+0.1)

image.plot((depth*-1), horizontal=F, smallplot= c(0.6, 0.63,0.1,0.9), col = depth.heat(1000), legend.only=TRUE,  axis.args = list(cex.axis = 1,
at = c(0, 20, 40, 60, 80, 100)+15, labels =rev(c("0","20","40","60","80","100")), 
tcl=0.3, lwd=0.5, hadj=0.5), axes=F, zlim=c(0, 115))

#dev.off()


mean(depth)

cellStats(depth, stat='sd')
```



## Mud model predictions
```{r}



mud.trim<- mud.r
mud.trim [mud.trim< 0.1] <- NA
#display.brewer.all()
prob.heat<- colorRampPalette(rev(brewer.pal(11,'Spectral')))



#pdf(file="F:/Seabed_map_paper/figures/4mud_pred.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
#plot(depth, add=T, col=depth.heat(1000), legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)

plot(mud.trim, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000) )

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=0.9)

plot(sub.spdfT[sub.spdfT$Substratum1 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)

#dev.off()

#plot(sub.pdfT3.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[1][sub.spdfT3.trim@data$Substratum3], cex=0.5)

#plot(sub.spdfT4.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[1][sub.spdfT4.trim@data$Substratum4], cex=2.4)

#pdf(file="F:/Seabed_map_paper/Figures/4mud_legend.pdf",width=2, height=2)

par(mar=c(0,0,0,0)+0.1)
plot(sub.spdfT, col='white')

legend('center', bg='white', legend=leg, fill=c( test.cols2[1], 'black', 'black', 'black', 'black', 'black'))


#dev.off()



```

#heat map legend
```{r}
#pdf(file="E:/SIFT_Jan2020/report_figs/substarat_leg.pdf",width=4.5, height=4.5) 

par(mar=c(0,0,0,0)+0.1)

image.plot((mud.r), horizontal=T, smallplot= c(0.1, 0.9,0.6,0.63), col = prob.heat(1000), legend.only=TRUE,  axis.args = list(cex.axis = 1,
at = c(0, 22, 50, 75, 100), labels =c("0","0.25","0.5","0.75","1.0"), 
tcl=0.3, lwd=0.5, hadj=0.5), axes=F, zlim=c(0, 100))

#dev.off()

```



## Mud RF prediction
```{r}


rf.cols<- c(brewer.pal(11,'Spectral')[11], brewer.pal(11,'Spectral')[1])

#pdf(file="F:/Seabed_map_paper/figures/Sup_mud_RF.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)

mud.rf.trim<- mud.rf
mud.rf.trim [mud.rf.trim<2 ] <- NA

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(mud.rf.trim, add=T, legend=FALSE, axes=FALSE, col=rf.cols)
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "mud" ,], pch=19, add=T, col=
      test.cols2[1], cex=1.1)

#plot(sub.pdfT3.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[3:6][sub.spdfT3.trim@data$Substratum3], cex=0.5)

#plot(sub.spdfT4.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[c(3,5,6)][sub.spdfT4.trim@data$Substratum4], cex=2.4)


#legend('bottomright', legend=leg, fill=c( test.cols2[1], 'black', 'black', 'black', 'black', 'black'))

#dev.off()

```

##mud krig
```{r}


mud.k.trim<- mud.k
mud.k.trim [mud.k.trim< 0.1] <- NA


#pdf(file="F:/Seabed_map_paper/figures/Sup_mud_K.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(mud.k.trim, add=T, legend=FALSE, axes=FALSE, col= prob.heat((1000)))
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "mud" ,], pch=19, add=T, col=
       test.cols2[1], cex=1.1)

#plot(sub.pdfT3.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[1][sub.spdfT3.trim@data$Substratum3], cex=0.5)

#plot(sub.spdfT4.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[1][sub.spdfT4.trim@data$Substratum4], cex=2.4)


#legend('bottomright', legend=leg, fill=c( test.cols2[1], 'black', 'black', 'black', 'black', 'black'))


#dev.off()

#hist(mud.k)

```



# Muddy_sand

## Muddy_sand model predictions
```{r}

# 
muddy_s.trim<- muddy_sand.r
muddy_s.trim [muddy_s.trim< 0.1] <- NA

#pdf(file="F:/Seabed_map_paper/figures/Sup_Sandy_mud_RF.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
#plot(depth, add=T, col=depth.heat(1000), legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)

plot(muddy_s.trim, col=prob.heat(1000), add=T, legend=FALSE, axes=FALSE)


plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

#dev.off()

#plot(sub.pdfT3.trim, pch=19, add=T, col=
 #      brewer.pal(8,"Dark2")[3:6][sub.spdfT3.trim@data$Substratum3], cex=0.5)

#plot(sub.spdfT4.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[c(3,5,6)][sub.spdfT4.trim@data$Substratum4], cex=2.4)


#pdf(file="F:/Seabed_map_paper/Figures/4muddy_s_legend.pdf",width=2, height=2)

# par(mar=c(0,0,0,0)+0.1)
# plot(sub.spdfT, col='white')
# 
# legend('center', bg='white', legend=leg, fill=c('1', test.cols2[2], '1', '1', '1','1'))

#dev.off()





```

## Muddy_sand RF prediction
```{r}
# 



muddy_s.rf.trim<- muddy_sand.rf
muddy_s.rf.trim [muddy_s.rf.trim<2 ] <- NA


#pdf(file="F:/Seabed_map_paper/figures/Sup_muddy_sand_RF.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)



plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(muddy_s.rf.trim, add=T, legend=FALSE, axes=FALSE, col=rf.cols)
plot(shore, col='black', add=T, lwd=1)


plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

#plot(sub.pdfT3.trim, pch=19, add=T, col=
 #      brewer.pal(8,"Dark2")[3:6][sub.spdfT3.trim@data$Substratum3], cex=0.5)

#plot(sub.spdfT4.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[c(3,5,6)][sub.spdfT4.trim@data$Substratum4], cex=2.4)

#legend('bottomright', legend=leg, fill=c('1', test.cols2[2], '1', '1', '1','1'))

#dev.off()

```


##muddy_sand krig
```{r}



muddy_s_k.trim<- muddy_sand.k
muddy_s_k.trim [muddy_s_k.trim< 0.1] <- NA


#pdf(file="F:/Seabed_map_paper/figures/Sup_muddy_sand_K.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(muddy_s_k.trim, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "muddy_sand" ,], pch=19, add=T, col=
       test.cols2[2], cex=1.1)

#plot(sub.pdfT3.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[1][sub.spdfT3.trim@data$Substratum3], cex=0.5)

#plot(sub.spdfT4.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[1][sub.spdfT4.trim@data$Substratum4], cex=2.4)


#legend('bottomright', legend=leg, fill=c('black', test.cols2[2], 'black', 'black', 'black', 'black'))


#dev.off()

#hist(muddy_sand.k)

```


# sand

## sand model predictions
```{r}



sand.trim<- sand.r
sand.trim [sand.trim< 0.1] <- NA



#pdf(file="F:/Seabed_map_paper/figures/4sand_pred_noD.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
#plot(depth, add=T, col=depth.heat(1000), legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)


plot(sand.trim, col = prob.heat(1000), add=T, legend=FALSE, axes=FALSE)


plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "sand" ,], pch=19, add=T, col=
       test.cols2[3], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

plot(sub.spdfT3.trim[sub.spdfT3.trim$Substratum3 == "sand" ,], pch=19, add=T, col=
      test.cols2[3], cex=1.1)

#dev.off()

#plot(sub.spdfT4.trim[sub.spdfT4.trim$Substratum4 == "sand" ,], pch=19, add=T, col=
     #  brewer.pal(8,"Dark2")[3], cex=1.1)


#pdf(file="F:/Seabed_map_paper/Figures/4sand_legend.pdf",width=2, height=2)

par(mar=c(0,0,0,0)+0.1)
plot(sub.spdfT, col='white')

legend('center', bg='white', legend=leg, fill=c('1', '1',  test.cols2[3], '1', '1', '1'))



#dev.off()


```

## sand RF prediction
```{r}



sand.rf.trim<- sand.rf
sand.rf.trim [sand.rf.trim<2 ] <- NA


#pdf(file="F:/Seabed_map_paper/figures/Sup_sand_RF.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)


plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(sand.rf.trim, add=T, legend=FALSE, axes=FALSE,col=rf.cols)
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

plot(sub.spdfT3.trim[sub.spdfT3.trim$Substratum3 == "sand" ,], pch=19, add=T, col=
      test.cols2[3], cex=1.1)



#legend('bottomright', legend=leg, fill=c('1', '1',  test.cols2[3], '1', '1', '1'))


#dev.off()

```


##sand krig
```{r}



sand_k.trim<- sand.k
sand_k.trim [sand_k.trim< 0.1] <- NA


#pdf(file="F:/Seabed_map_paper/figures/Sup_sand_K.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(sand_k.trim, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "sand" ,], pch=19, add=T, col=
        test.cols2[3], cex=1.1)

#plot(sub.pdfT3.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[1][sub.spdfT3.trim@data$Substratum3], cex=0.5)

#plot(sub.spdfT4.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[1][sub.spdfT4.trim@data$Substratum4], cex=2.4)


#legend('bottomright', legend=leg, fill=c('black', 'black',  test.cols2[3], 'black', 'black', 'black'))


#dev.off()


```



# gravel


## gravel model predictions
```{r}




gravel.trim<- gravel.r
gravel.trim [gravel.trim< 0.1] <- NA


#pdf(file="F:/Seabed_map_paper/figures/4gravel_pred_noD.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
#plot(depth, add=T, col=depth.heat(1000), legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)


plot(gravel.trim, col = prob.heat(1000), add=T, legend=FALSE, axes=FALSE)


plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

#dev.off()

#plot(sub.spdfT3.trim[sub.spdfT3.trim$Substratum3 == "gravel" ,], pch=19, add=T, col=
  #   brewer.pal(8,"Dark2")[4], cex=1.1)

#plot(sub.spdfT4.trim[sub.spdfT4.trim$Substratum4 == "sand" ,], pch=19, add=T, col=
 #      brewer.pal(8,"Dark2")[4], cex=1.1)


#pdf(file="F:/Seabed_map_paper/Figures/4gravel_legend.pdf",width=2, height=2)

par(mar=c(0,0,0,0)+0.1)
plot(sub.spdfT, col='white')


legend('center',bg='white', legend=leg, fill=c('1','1','1', test.cols2[4], '1', '1'))

#dev.off()


```

## gravel RF prediction
```{r}
 

gravel.rf.trim<- gravel.rf
gravel.rf.trim [gravel.rf.trim<2 ] <- NA

#pdf(file="F:/Seabed_map_paper/figures/Sup_gravel_RF.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)


plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(gravel.rf.trim, add=T, legend=FALSE, axes=FALSE, col=rf.cols)
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

#plot(sub.spdfT3.trim[sub.spdfT3.trim$Substratum3 == "gravel" ,], pch=19, add=T, col=
#     brewer.pal(8,"Dark2")[5], cex=1.1)

#plot(sub.spdfT4.trim[sub.spdfT4.trim$Substratum4 == "sand" ,], pch=19, add=T, col=
    #   brewer.pal(8,"Dark2")[4], cex=1.1)



#legend('bottomright', legend=leg, fill=c('1','1','1', test.cols2[4], '1', '1'))

#dev.off()

```


##gravel krig
```{r}


gravel.k.trim<- gravel.k
gravel.k.trim [gravel.k.trim< 0.1] <- NA

#pdf(file="F:/Seabed_map_paper/figures/Sup_gravel_K.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(gravel.k.trim, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "gravel" ,], pch=19, add=T, col=
       test.cols2[4], cex=1.1)

#plot(sub.pdfT3.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[1][sub.spdfT3.trim@data$Substratum3], cex=0.5)

#plot(sub.spdfT4.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[1][sub.spdfT4.trim@data$Substratum4], cex=2.4)


#legend('bottomright', legend=leg, fill=c('black', 'black', 'black', test.cols2[4], 'black', 'black'))


#dev.off()

```






#Cobble-boulder joint


## Cob_boulder model predictions
```{r}


cob_boulder.trim<- cob_boulder.r
cob_boulder.trim [cob_boulder.trim< 0.1] <- NA



#pdf(file="F:/Seabed_map_paper/figures/4cob_boulder_pred_noD.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
#plot(depth, add=T, col=depth.heat(1000), legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)



plot(cob_boulder.trim, col = prob.heat(1000), add=T, legend=FALSE, axes=FALSE)


plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)


plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)

plot(sub.spdfT3.trim[sub.spdfT3.trim$Substratum3 == "cob_boulder" ,], pch=19, add=T, col=
     test.cols2[5], cex=1.1)

#dev.off()

# plot(sub.spdfT4.trim[sub.spdfT4.trim$Substratum4 == "boulder" ,], pch=19, add=T, col=
#        brewer.pal(8,"Dark2")[6], cex=1.1)
# 


#pdf(file="F:/Seabed_map_paper/Figures/4cob_boulder_legend.pdf",width=2, height=2)

par(mar=c(0,0,0,0)+0.1)
plot(sub.spdfT, col='white')

legend('center',bg='white', legend=leg, fill=c('1','1','1','1', test.cols2[5],'1'))

#dev.off()




```

## Cob-boulder RF prediction
```{r}



cob_boulder.rf.trim<- cob_boulder.rf
cob_boulder.rf.trim [cob_boulder.rf.trim<2 ] <- NA


#pdf(file="F:/Seabed_map_paper/figures/Sup_cob_boulder_RF.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)


plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(cob_boulder.rf.trim, add=T, legend=FALSE, axes=FALSE, col=rf.cols)
plot(shore, col='black', add=T, lwd=1)


plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)

plot(sub.spdfT3.trim[sub.spdfT3.trim$Substratum3 == "cob_boulder" ,], pch=19, add=T, col=
     test.cols2[5], cex=1.1)



#legend('bottomright', legend=leg, fill=c('1','1','1','1', test.cols2[5],'1'))

#dev.off()

```


##Cob_boulder krig
```{r}

 

cob_boulder.k.trim<- cob_boulder.k
cob_boulder.k.trim [cob_boulder.k.trim< 0.1] <- NA

#pdf(file="F:/Seabed_map_paper/figures/Sup_cob_boulder_K.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(cob_boulder.k.trim, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "cob_boulder" ,], pch=19, add=T, col=
       test.cols2[5], cex=1.1)

#plot(sub.pdfT3.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[1][sub.spdfT3.trim@data$Substratum3], cex=0.5)

#plot(sub.spdfT4.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[1][sub.spdfT4.trim@data$Substratum4], cex=2.4)


#legend('bottomright', legend=leg, fill=c( 'black', 'black', 'black', 'black', test.cols2[5], 'black'))


#dev.off()

#hist(cob_boulder.k)

```


# rock


## rock model predictions
```{r}



rock.trim<- rock.r
rock.trim [rock.trim< 0.1] <- NA



#pdf(file="F:/Seabed_map_paper/figures/4rock_pred_noD.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
#plot(depth, add=T, col=depth.heat(1000), legend=FALSE, axes=FALSE)
plot(shore, col='black', add=T, lwd=1)


plot(rock.trim, col = prob.heat(1000), add=T, legend=FALSE, axes=FALSE)


plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)



#dev.off()
#plot(sub.spdfT3.trim[sub.spdfT3.trim$Substratum3 == "rock" ,], pch=19, add=T, col=
 #    brewer.pal(8,"Dark2")[6], cex=1.1)

#plot(sub.spdfT4.trim[sub.spdfT4.trim$Substratum4 == "rock" ,], pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[7], cex=1.1)


#pdf(file="F:/Seabed_map_paper/Figures/4rock_legend.pdf",width=2, height=2)

par(mar=c(0,0,0,0)+0.1)
plot(sub.spdfT, col='white')


legend('center', bg='white', legend=leg, fill= c('1','1','1','1','1', test.cols2[6]))

#dev.off()



```

## rock RF prediction
```{r}

rock.rf.trim<- rock.rf
rock.rf.trim [rock.rf.trim<2 ] <- NA


#pdf(file="F:/Seabed_map_paper/figures/Sup_rock_RF.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)


plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(rock.rf.trim, add=T, legend=FALSE, axes=FALSE, col=rf.cols)
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)

#plot(sub.spdfT3.trim[sub.spdfT3.trim$Substratum3 == "rock" ,], pch=19, add=T, col=
 #    brewer.pal(8,"Dark2")[6], cex=1.1)

#plot(sub.spdfT4.trim[sub.spdfT4.trim$Substratum4 == "rock" ,], pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[7], cex=1.1)

#legend('bottomright', legend=leg, fill= c('1','1','1','1','1', test.cols2[6]))

#dev.off()

```


##rock krig
```{r}

 
rock.k.trim<- rock.k
rock.k.trim [rock.k.trim< 0.1] <- NA

#pdf(file="F:/Seabed_map_paper/figures/Sup_rock_K.pdf",width=6, height=6)
par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F, add=T)
plot(rock.k.trim, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000))
plot(shore, col='black', add=T, lwd=1)

plot(sub.spdfT, pch=19, add=T, col=
       'black', cex=1)

plot(sub.spdfT[sub.spdfT$Substratum1 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)

plot(sub.spdfT2.trim[sub.spdfT2.trim$Substratum2 == "rock" ,], pch=19, add=T, col=
       test.cols2[6], cex=1.1)

#plot(sub.pdfT3.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[1][sub.spdfT3.trim@data$Substratum3], cex=0.5)

#plot(sub.spdfT4.trim, pch=19, add=T, col=
#       brewer.pal(8,"Dark2")[1][sub.spdfT4.trim@data$Substratum4], cex=2.4)


#legend('bottomright', legend=leg, fill=c('black', 'black', 'black', 'black', 'black', test.cols2[6]))


#dev.off()

#hist(rock.k)


```







# model plots legend
```{r}

#pdf(file="F:/SIFT2021/May on/report/preds_legend.pdf",width=4.5, height=4.5)

image.plot((mud.r), horizontal=T, smallplot= c(0.1, 0.9,0.6,0.63), col = rev(terrain.colors(1000)), legend.only=TRUE,  axis.args = list(cex.axis = 1,
at = c(0, 50, 100), labels =c("0","0.5","1.0"), 
tcl=0.3, lwd=0.5, hadj=0.5), axes=F, zlim=c(0, 100))

#dev.off()

```


## combined map
```{r}
# make a set of colour pallets for mud, muddy_sand, sand, gravel, cob_boulder and rock...so 6 in all.

# Trim each raster to only contain values of reasonable magnitude
density(mud.r)
density(muddy_sand.r)
density(sand.r)
density(gravel.r)
density(cob_boulder.r)
density(rock.r)

# write files for later use
#writeRaster(mud.r, 'F:/SIFT2021/May on/GIS_data/analysis_files/mud_pred.grd', format='raster')

#writeRaster(muddy_sand.r, 'F:/SIFT2021/May on/GIS_data/analysis_files/muddy_s_pred.grd', format='raster')


#writeRaster(sand.r, 'F:/SIFT2021/May on/GIS_data/analysis_files/sand_pred.grd', format='raster')


#writeRaster(gravel.r, 'F:/SIFT2021/May on/GIS_data/analysis_files/gravel_pred.grd', format='raster')

#writeRaster(cob_boulder.r, 'F:/SIFT2021/May on/GIS_data/analysis_files/cob_boul_pred.grd', format='raster')

#writeRaster(rock.r, 'F:/SIFT2021/May on/GIS_data/analysis_files/rock_pred.grd', format='raster')


mud.trim<- mud.r
muddy_sand.trim<- muddy_sand.r
sand.trim<- sand.r
gravel.trim<- gravel.r
cob_boulder.trim<- cob_boulder.r
rock.trim<- rock.r


#density(mud.r)
mud.trim [mud.trim< 0.51] <- NA
#density(mud.trim)



#density(muddy_sand.r)
muddy_sand.trim [muddy_sand.trim< 0.51] <- NA
#density(muddy_sand.trim)


#density(sand.r)
sand.trim [sand.trim< 0.11] <- NA
#density(sand.trim)

#density (gravel.r)
gravel.trim [gravel.trim< 0.05] <- NA
#density(gravel.trim)

#density(cob_boulder.r)
cob_boulder.trim [cob_boulder.trim< 0.11] <- NA
#density(cob_boulder.trim)

#density(rock.r)
rock.trim [rock.trim< 0.51] <- NA
#density(rock.trim)


# ##################################### all at 0.1
# 
# #density(mud.r)
# mud.trim [mud.trim< 0.1] <- NA
# #density(mud.trim)
# 
# 
# 
# #density(muddy_sand.r)
# muddy_sand.trim [muddy_sand.trim< 0.1] <- NA
# #density(muddy_sand.trim)
# 
# 
# #density(sand.r)
# sand.trim [sand.trim< 0.1] <- NA
# #density(sand.trim)
# 
# #density (gravel.r)
# gravel.trim [gravel.trim< 0.1] <- NA
# #density(gravel.trim)
# 
# #density(cob_boulder.r)
# cob_boulder.trim [cob_boulder.trim< 0.1] <- NA
# #density(cob_boulder.trim)
# 
# #density(rock.r)
# rock.trim [rock.trim< 0.1] <- NA
# #density(rock.trim)







#display.brewer.all()
# mud.heat<- colorRampPalette(brewer.pal(9,'YlOrRd')[c(1,4:8)], alpha=TRUE)
# muddy_s.heat<- colorRampPalette(brewer.pal(9,'Purples')[c(1,4:8)], alpha=TRUE)
# sand.heat<- colorRampPalette(brewer.pal(9,'YlOrRd')[c(1:4)], alpha=TRUE, bias=2)
# gravel.heat<- colorRampPalette(brewer.pal(9,'Greys')[c(5:9)], alpha=TRUE)
# cob_boul.heat<- colorRampPalette(brewer.pal(9,'Greens')[c(1,5:9)], alpha=TRUE)
# rock.heat<- colorRampPalette(brewer.pal(9,'Blues')[c(1,5:9)], alpha=TRUE)


# addalpha()
addalpha <- function(colors, alpha=1.0) {
  r <- col2rgb(colors, alpha=T)
  # Apply alpha
  r[4,] <- alpha*255
  r <- r/255.0
  return(rgb(r[1,], r[2,], r[3,], r[4,]))
}

#### maybe use this to go from test cols 1 to white to the shade required 1st try just the colour with an alpha value

mud.alt <- colorRampPalette(c('wheat', test.cols[1], test.cols2[1]), alpha=TRUE, bias=1)

mud.s.alt <- colorRampPalette(c('darkseagreen1', test.cols[2], test.cols2[2]), alpha=TRUE, bias=1)
sand.alt <- colorRampPalette(c('lightgoldenrodyellow', test.cols[3], test.cols2[3]), alpha=TRUE, bias=1.5)
gravel.alt <- colorRampPalette(c('lightpink1', test.cols[4], test.cols2[4]), alpha=TRUE, bias=0.3)
cob_bol.alt<- colorRampPalette(c('thistle1', test.cols[5], test.cols2[5]), alpha=TRUE, bias=1.5)
rock.alt<- colorRampPalette(c('lightsteelblue2', test.cols[6], test.cols2[6]), alpha=TRUE, bias=1)



 mud.alt.single <- addalpha(test.cols2[1], alpha=0.7)
 mud.s.alt.single <- addalpha(test.cols2[2], alpha=0.7)
 sand.alt.single <- addalpha(test.cols2[3], alpha=0.7)
 gravel.alt.single <-addalpha(test.cols2[4], alpha=0.7)
 cob_bol.alt.single<- addalpha(test.cols2[5], alpha=0.7)
 rock.alt.single<- addalpha(test.cols2[6], alpha=0.7)


#pdf(file="F:/Seabed_map_paper/figures/5cobined_substarta_map_01.pdf",width=6, height=6)

par(mar=c(0,0,0,0)+0.1)

plot(sub.spdfT, col='white')
plot(hill.trim, col=hill.col(1000), legend=F,  axes=FALSE, box=F, add=T)
#plot(depth, add=T, col=depth.heat(1000), legend=FALSE, axes=FALSE)

plot(cob_boulder.trim, add=T, legend=FALSE, axes=FALSE, col=cob_bol.alt(1000), alpha=0.7)

plot(gravel.trim, add=T, legend=FALSE, axes=FALSE, col=gravel.alt(1000), alpha=0.4)

plot(sand.trim, add=T, legend=FALSE, axes=FALSE, col=sand.alt(1000), alpha=0.6)


plot(muddy_sand.trim, add=T, legend=FALSE, axes=FALSE, col=mud.s.alt(1000), alpha=0.5)


plot(rock.trim, add=T, legend=FALSE, axes=FALSE, col=rock.alt(1000), alpha=0.8)


plot(mud.trim, add=T, legend=FALSE, axes=FALSE, col=mud.alt(1000), alpha=0.6)

plot(shore, col='black', add=T, lwd=1)


#dev.off()


 leg.cols<- c(mud.alt.single, mud.s.alt.single, sand.alt.single,
gravel.alt.single, cob_bol.alt.single, rock.alt.single)


#pdf(file="F:/Seabed_map_paper/Figures/5combined_legend.pdf",width=2, height=2)

par(mar=c(0,0,0,0)+0.1)
plot(sub.spdfT, col='white')

legend('center', bg='white', legend=leg, fill= leg.cols)



#dev.off()

#plot(shore, col='black', add=T, lwd=1)

# leg.cols<- c(brewer.pal(9,'YlOrRd')[7],  brewer.pal(9,'Purples')[7], brewer.pal(9,'YlOrRd')[4], brewer.pal(9,'Greys')[4], brewer.pal(9,'Greens')[8],brewer.pal(9,'Blues')[7])
# 
# alt.cols<-brewer.pal(8,"Dark2")[c(2:6,8)]
# 
# legend('bottomright', legend=leg, fill= alt.cols)

#dev.off()

```


##
## area calcs-  wider area combined map
```{r}
# make a set of colour pallets for mud, muddy_sand, sand, gravel, cob_boulder and rock...so 6 in all.

# Trim each raster to only contain values of specific magnitude
density(mud.r)
density(muddy_sand.r)
density(sand.r)
density(gravel.r)
density(cob_boulder.r)
density(rock.r)

mud.trim<- mud.r
muddy_sand.trim<- muddy_sand.r
sand.trim<- sand.r
gravel.trim<- gravel.r
cob_boulder.trim<- cob_boulder.r
rock.trim<- rock.r


hist(mud.r)
mud.trim [mud.trim<= 0.1] <- NA
density(mud.trim)

# overall area
plot(depth, colNA="dodgerblue")
cell.No<-cellStats(depth, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(depth)[1]*res(depth)[2])
round(raster_area/(1000*1000), 2)

###area calc
plot(mud.trim, colNA="dodgerblue")

cell.No<-cellStats(mud.trim, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(mud.trim)[1]*res(mud.trim)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)

#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

density(muddy_sand.r)
muddy_sand.trim [muddy_sand.trim<= 0.1] <- NA
density(muddy_sand.trim)

###area calc
plot(muddy_sand.trim, colNA="dodgerblue")

cell.No<-cellStats(muddy_sand.trim, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(muddy_sand.trim)[1]*res(muddy_sand.trim)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)

#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

density(sand.r)
sand.trim [sand.trim<= 0.1] <- NA
density(sand.trim)

###area calc
plot(sand.trim, colNA="dodgerblue")

cell.No<-cellStats(sand.trim, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(sand.trim)[1]*res(sand.trim)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)
#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)



density (gravel.r)
gravel.trim [gravel.trim <= 0.1] <- NA
density(gravel.trim)

###area calc
plot(gravel.trim, colNA="dodgerblue")

cell.No<-cellStats(gravel.trim, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(gravel.trim)[1]*res(gravel.trim)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)
#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

density(cob_boulder.r)
cob_boulder.trim [cob_boulder.trim<= 0.1] <- NA
density(cob_boulder.trim)

###area calc
plot(cob_boulder.trim, colNA="dodgerblue")

cell.No<-cellStats(cob_boulder.trim, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(cob_boulder.trim)[1]*res(cob_boulder.trim)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)
#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

density(rock.r)
rock.trim [rock.trim<= 0.1] <- NA
density(rock.trim)

###area calc
plot(rock.trim, colNA="dodgerblue")

cell.No<-cellStats(rock.trim, function(i, ...) sum(!is.na(i)))

raster_area<- cell.No * (res(rock.trim)[1]*res(rock.trim)[2])

#area at >0.1
round(raster_area/(1000*1000), 2)
#proportion
round(round(raster_area/(1000*1000), 2)/total.area,3)

```
